---
title: "ch9_hm_svm"
author: "Andrey Glushko"
date: "13 March 2016"
output: html_document
---

In this problem, you will use simulation to evaluate (by Monte Carlo) the expected misclassification error rate given a particular generating model. Let yi be equally divided between classes 0 and 1, and let xi is R^10 be normally distributed.

Preparing train data:

```{r}
library(MASS)
set.seed(10111)

gen1.sample = function(n1,n2){
  top = matrix(rnorm(500), ncol = 10)
  bl = matrix(rnorm(250,mean = 1),ncol=5)
  br = matrix(rnorm(250),ncol=5)
  bottom = cbind(bl,br)
  full = rbind(top,bottom)
  full = as.data.frame(full)
  full$y = as.factor(rep(0:1, each=50))
  return(full)
}

gen.sample=function(n1, n2) {
  x1=mvrnorm(n=n1, mu=rep(0,10), Sigma=diag(10))
  x2=mvrnorm(n=n2, mu=rep(c(1,0),c(5,5)), Sigma=diag(10))
  x=rbind(x1,x2);
  x=matrix(x, nrow=100, ncol=10)
  y=rep(c(0,1), c(n1,n2))
  y=matrix(y,n1+n2,1)
  sample=data.frame(x,y=as.factor(y))
  return(sample)
}

train.dat=gen1.sample(50, 50)

#x1=mvrnorm(50, mu=rep(0,10), Sigma=diag(10))
#x2=mvrnorm(50, mu=c(1,1,1,1,1,0,0,0,0,0), Sigma=diag(10))
#x=rbind(x1,x2);
#dim(x)
#x=matrix(x, nrow=100, ncol=10)
#dim(x)
#xt = x[51:100,4]
#hist(xt)
#mean(xt)

#y=rep(c(0,1), c(50,50))
#y=matrix(y,100,1)
#train.data=data.frame(x,y)
#train.dat=data.frame(x,y=as.factor(y))
```

Train SVM using train data:
```{r}
library(e1071)

num_it = 100;
svm.error=rep(0,num_it)

for(i in 1:num_it) {
  # train
  train.dat=gen.sample(50, 50)
  svmfit=svm(train.dat$y~., data=train.dat)
  
  # test
  test.dat=gen.sample(100, 100)
#  test.dat=gen.sample(50, 50)
  svmpred = predict(svmfit, test.dat)
  
  # test error
#  svm.error[i] = (sum(svmpred != test.dat$y)) / length(test.dat$y)  
  svm.error[i] = mean(svmpred != test.dat$y)
}

# average error:
err = sum(svm.error) / length(svm.error)
err


# svm + cv:
#train.dat=gen.sample(50, 50)
#svmfit=svm(train.dat$y~., data=train.dat)
#tune.out = tune(svm, y~., data=train.dat)
#summary(tune.out)
```

Train SVM using kernel="linear" train data:
```{r}
svm.error=rep(0,num_it)

for(i in 1:num_it) {
  # train
  train.dat=gen.sample(50, 50)
  svmfit=svm(train.dat$y~.,kernel="linear",data=train.dat)
  
  # test
  test.dat=gen.sample(100, 100)
  svmpred = predict(svmfit, test.dat)
  
  # test error
  svm.error[i] = mean(svmpred != test.dat$y)
}

# average error:
err = sum(svm.error) / length(svm.error)
err
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
